# Harmful events and their economic consequences caused by Storm and severe weather events in the US based on NOAA Storm Database(Coursera repdata-002 Assigment 2)  
----
----


The analysis based on Storm and severe events that have the biggest health and economic consequences in the US.
We can say after the analysis the most harmful events in the US are caused by tornados. In tornados **96979** people died or injured and **$ 3212258.2 million** property damage caused.

 
----
----

## Data Processing
1. Loading repdata-data-StormData.csv.bz2
2. Filter and aggregation
  * Aggregate events by Event types and sum the number of deaths and injuries
  * Aggregate events by Event types and sum the property damage variable
3. Sort the results by the sum variable descending

----
Loading ggplot2 for plots
```{r}
library(ggplot2)
```


Loading the data
```{r cache=TRUE}
stormData <- read.csv(bzfile("data/repdata-data-StormData.csv.bz2"))
```
## Summary of Deaths in events
```{r}
summary(stormData$FATALITIES)
```
## Summary of Injuries in events
```{r}
summary(stormData$INJURIES)
```
## Summary of Property damage in events
```{r}
summary(stormData$PROPDMG)
```
```{r}
#Aggregation by event type and sum deaths and injuries
harmfulEvents <-aggregate(FATALITIES+INJURIES ~ EVTYPE, data=stormData, FUN=sum)
#more talkable col names
colnames(harmfulEvents) <- c("EventType","Sum")
#Sorting and top10
harmfulEvents.summary <-head(harmfulEvents[order(-harmfulEvents$Sum),],10)
```
## Results for Top 10 harmful event types in the US(Deaths + Injuries)
```{r}
print(harmfulEvents.summary)
```
```{r}
#harmful events plot
ggplot(data=harmfulEvents.summary, aes(x=EventType, y=Sum, fill=EventType)) + geom_bar(stat="identity") + ylab("Sum of deaths and injuries") + xlab("Event") +
  ggtitle("Most harmful events in the US")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#   
   
----
   
# 
   
```{r}
#Aggregation by event type and sum property damage
harmfulEventsDamage <-aggregate(PROPDMG ~ EVTYPE, data=stormData, FUN=sum)
#more talkable col names
colnames(harmfulEventsDamage) <- c("EventType","SumOfDamage")
#Sorting and top10
harmfulEventsDamage.summary <-head(harmfulEventsDamage[order(-harmfulEventsDamage$SumOfDamage),],10)
```
## Results for Top 10 harmful event types what caused the greatest economic consequences in the US
```{r}
print(harmfulEventsDamage.summary)
```
```{r}
#economic consequences plot
ggplot(data=harmfulEventsDamage.summary, aes(x=EventType, y=SumOfDamage, fill=EventType)) + geom_bar(stat="identity") + ylab("Sum of property damage in $ million") + xlab("Event") +
  ggtitle("Greatest economic consequences of events")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


